Performance comparison of Dask and Apache Spark on HPC systems

### 2022-08-17 latex/paper.tex
[REMOVE] highlight from previous revision.

\abstract[Summary]{
	The general increase in data size and data sharing motivates the adoption of Big Data strategies in a number of scientific disciplines.
	However, while several options are available, no particular guidelines exist for the selection of a Big Data engine.
	In this paper, we compare the runtime performance of two popular Big Data engines with
	Python APIs, Apache Spark, and Dask, in processing neuroimaging pipelines.
	Our experiments use three synthetic data-intensive applications to process
	the \SI{606}{\gibi\byte} BigBrain image and an actual pipeline to process data
	from thousands of anatomical images.
	We benchmark these applications on a dedicated HPC cluster running the Lustre file system while using varying
	combinations of number of nodes, file size, and task duration.
	Our results show that although there are slight differences between Dask and Spark,
	the performance of the engines is comparable for data-intensive applications.
	However, Spark requires more memory than Dask, which can lead to 
	slower runtime depending on configuration and infrastructure.
	In general, the limiting factor was the data transfer time.
	While both engines are suitable for neuroimaging, more efforts need to be put
	to reduce the data transfer time and the memory footprint of applications.
}

@@

\section{Introduction}
The rise in data sharing coupled with improved data collection technologies leads neuroimaging into the Big Data era\cite{ALFAROALMAGRO2018400, van2014human, ConpPortal}.
While standard neuroimaging workflow engines, such as Nipype\cite{Nipype:11}, are well rounded to process standard compute-intensive pipelines,
they lack Big Data strategies (i.e., in-memory computing, data locality, and lazy-evaluation) to improve the performance of increasingly prevalent data-intensive pipelines.
A previous study\cite{8752675} noted the importance of in-memory computing and data locality to improve the performance of data-intensive pipelines.
In this work, we study how performance benchmarks generalize across Big Data engines in an HPC environment.
In particular, we study in-depth the performance difference between Dask\cite{Dask:15} and Apache Spark\cite{Spark:16} for their suitability to process neuroimaging pipelines.
Our objective is to assess whether Spark or Dask has substantial performance benefits in processing data-intensive HPC pipelines.

Spark and Dask both provide in-memory computing, data locality, and lazy-evaluation, typical for Big Data engines.
The scheduler for both engines operates dynamically, which benefits applications with runtime unknown ahead of time\cite{Dask:15}.
They also provide rich high-level API and support a variety of schedulers and deployment environments, such as Mesos\cite{hindman2011mesos}, YARN\cite{vavilapalli2013apache}, Kubernetes, and HPC clusters.
Although sharing similarities, these engines have differences.

First and foremost, Spark is written in Scala while Dask is in Python.
The popularity of Python in the scientific community arguably gives an edge to Dask due to the serialization cost between Python and Scala.
However, if not careful, the Python Global Interpreter Lock (GIL) can significantly reduce the parallelism of an application in some cases.
The difference in programming languages also provides different benefits.
On the one hand, as part of the Scipy ecosystem, Dask provides almost transparent integration with APIs such as Numpy arrays, Pandas dataframe, or RAPIDS GPU accelerated code framework.
On the other hand, Spark's Java, Scala, R, and Python APIs efficiently parallelize pipelines with minimal performance loss, due to the Java Virtual Machine.
Our study is restrained to performance, although we recognize that other factors affect the choice of an engine in practice.

While laptops or workstations can be sufficient for some applications, data-intensive neuroimaging pipelines often require large infrastructures.
This paper focuses on the performance of Big Data engines in HPCs environments.
We installed the Slurm scheduler and Lustre file system~\cite{Braam2019TheLS} on a dedicated cluster to mimic a typical HPC system.
\HL{Lustre is a reference file system for HPC environments, used in over 60 of the TOP100 supercomputers~{\cite{OpenSFS-lustre}}, offering parallel file accesses.}

In a previous study\cite{8943502} we found that there was no significant performance difference between Apache Spark and Dask for data-intensive applications.
However, using the network file system (NFS) led to an I/O bottleneck limiting application performance.
Another study~\cite{8588652} also reported network and I/O bottlenecks when concurrency increased.
This same study found that the startup overhead time for Spark was more prominent than for Dask.
On the contrary, the work in~\cite{Mehta:17} claims that the startup overhead is more significant for Dask than for Spark.
This contradiction suggests that the startup overhead might vary depending on cluster configuration.
The latter study processed a \SI{100}{\giga\byte} dataset using a neuroimaging pipeline and found Dask to be up to 14\% faster than Spark due to ``more efficient pipelining" and serialization time to Python.
On the contrary, the work in~\cite{10.1145/3225058.3225128} shows that with large datasets, Spark provides better speedup than Dask.
However, that study used the Dask Bag API, which has performance limitations compared to other Dask APIs.

The above studies compared the performance of the engine at a high level.
In comparison, our study provides a detailed analysis of the performance differences
and similarities\HL{, for HPC,} between Dask and Spark for data-intensive applications.

@@

\section{Background}
\subsection{Lustre}
The Lustre file system is a parallel file system \HL{of reference in} HPC environments.
Lustre offers a POSIX-compliant interface with the file system.
Installed in some of the most powerful supercomputers, the design of Lustre allows
scaling to thousands of nodes, petabytes of storage, and hundreds of gigabits per second of I/O throughput.

At the core of Lustre are the \textit{management servers} (MGSs) and \textit{metatdata server} (MDSs) in charge of the file operations and metadata,
the \textit{object storage targets} (OSTs) to store file data, and the \textit{object sotrage server} (OSSs) to handle the OSTs.
Lustre clients make requests to the MGSs to access files on the OSTs.
To handle failures in the system, the MGSs, MDSs, and OSTs have failover capabilities.
InfiniBand, Ethernet, or a combination of both can be used to interconnect these different components.

\HL{While other filesystem such as HDFS or cloud storage could be considered, we chose Lustre to best represent an HPC environment.
	HDFS makes use of permanent local storage on compute node, which is usually not feasable in HPC.
Morevoer, HPC node rarely have access to internet, limiting the use of cloud storage.}

\subsection{Dask}
Dask is a Python-based Big Data engine with growing popularity in the scientific Python ecosystem.
Dask was designed with data locality and in-memory computing in mind to mitigate the data transfer bottleneck in Big Data workflows.
Data locality, popularized by Map-Reduce~\cite{dean2008mapreduce}, schedules tasks where the data reside.
In-memory computing minimizes the overhead of transferring data to disk by keeping data in memory when possible.
Dask uses lazy evaluation to reduce unnecessary communication and computation.
The engine builds a dynamic graph before execution, allowing it to determine which task to compute.
Dask workflows can further reduce data transfers by leveraging multithreading whenever the Python GIL does not restrict it.
It achieves fault tolerance by recording data lineage, the sequence of operations used to modify the initial data.

Dask offers five data structures:
\href{https://docs.dask.org/en/latest/array.html}{Array},
\href{https://docs.dask.org/en/latest/bag.html}{Bag},
\href{https://docs.dask.org/en/latest/dataframe.html}{DataFrame},
\href{https://docs.dask.org/en/latest/delayed.html}{Delayed},
and \href{https://docs.dask.org/en/latest/futures.html}{Futures}.
Arrays offer a clone of NumPy API for distributed processing of large arrays.
Bags are a distributed collection of Python object that offers a programming abstraction similar to \href{https://toolz.readthedocs.io/en/latest/}{PyToolz}.
Dataframes are a parallel composition of \href{https://pandas.pydata.org/}{Pandas} Dataframes used to process a large amount of tabular data.
Dask Delayed offers an API for distributing arbitrary functions that do not fit in the above frameworks.
Lastly, Dask Futures can also execute arbitrary functions; however, it launches computation immediately rather than lazily.
Dask allows users to install only required components making it lightweight.

@@

The Dask engine is compatible with multiple distributed schedulers, including YARN and Mesos.
It also provides its own scheduler, \textit{Dask Distributed scheduler}.
We chose to use Dask Distributed scheduler to keep the environments similar between Dask and Spark.

@@

In the Spark Standalone scheduler, a \textit{leader} (a.k.a master) coordinates the resource provisioned by \textit{workers} in the cluster.
A \textit{driver}  process receives jobs from clients and requests workers from the leader.
Jobs are divided into stages to execute onto workers.
Each operation in a stage is represents by a high-level task in the computation graph.
Like Dask, Spark Standalone scheduler uses a LIFO policy to schedule tasks.
Spark Standalone has two execution modes: (1) the client mode, where the driver process runs in a dedicated process,
and (2) the cluster mode, where the driver runs within a worker process.
Our experiments use the client mode since cluster mode is not available in PySpark.

@@

Python is a standard programming language in the scientific community, offering numerous data processing libraries.
While serialization from Python to Java, an operation required when using Spark's Python API, creates overhead, we found it minimal in a previous study~\cite{8943502}.
We focus on PySpark API to have a similar environment between the different engines and due to its suitability for neuroimaging.

\section{Methods}
\subsection{Infrastructure}
We used the ``slashbin" cluster at Concordia University.
The cluster has eight compute nodes, four storage nodes, one login node, and one Lustre metadata node.
Each compute node has two 16-core Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz,
\SI{250}{\gibi\byte} of RAM, \SI{126}{\gibi\byte} of tmpfs,
% 6 $\times$ SSDs of \SI{447}{\gibi\byte} each with the XFS file system (no RAID enabled),
CentOS~8.1 and Linux kernel \textit{4.18.0-240.1.1.el8\_lustre.x86\_64}.
\HL{The compute and storage nodes are inter-connected by {\SI{10}{\giga\bit\per\second}} Ethernet link through a switch.
	The login node, hosting an NFS, is also connected by a {\SI{10}{\giga\bit\per\second}} Ethernet link to the compute nodes.}

We configured Lustre to have 1 MGS and 4 OSS.
Each OSS has 11 OSTs (disks) of \SI{8.8}{\tebi\byte}.
The OSTs are configured to have a maximum of \SI{1000}{\mebi\byte} of dirty data written into the client page cache.
Both the MDT and OSTs are configured to a maximum of 64 concurrent RPCs in flight.
A timeout of \SI{100}{\second} is set for Lustre, and OSTs checksum are disabled.
Lustre \textit{2.14.0-1.el8} is installed on both the Lustre clients and servers.
\HL{We used the default inode configuration for Lustre.}

\HL{Our system uses similar hardware as Compute Canada Beluga, and replicates as
	close as possible it's configuration.
	This let us make a fair comparison of Spark and Dask in the condition of a realistic
HPC system.}

Both Spark and Dask are configured to have eight worker processes, each with eight threads.
Each worker was allocated \SI{31.5}{\gibi\byte} of memory.
For Spark, the JVM heap space uses 10\% of workers' memory, and the location of the log is set on an NFS-mounted directory.
The worker workspace for both engines is located on Lustre due to the limited fast storage space on the worker nodes.
Other Dask configurations are left to default.
The Spark network timeout is set to \SI{600}{\second} and the executor heartbeat to \SI{120}{\second}.
A new Dask or Spark cluster was started and tore down for each experiment.
Dask \textit{2021.7.0} and Spark \textit{3.1.2} were used.

\subsection{Dataset}
We used BigBrain\cite{Amunts:13}, a 3-D image of the human brain with voxel
intensities ranging from 0 to 65,535. We converted the blocks into the
NIfTI format, a popular format in neuroimaging. We left the NIfTI blocks
uncompressed, resulting in total data size of \SI{606}{\gibi\byte}. To
evaluate the effect of block size, we split the image into 1000, 2500,
and 5000 files of \SI{606}{\mebi\byte}, \SI{242.4}{\mebi\byte}, and
\SI{121.2}{\mebi\byte}, respectively.
We used the \href{https://github.com/big-data-lab-team/sam}{sam} library to split the image.

@@

\subsection{Applications}
In this section, we describe the applications used to benchmark Dask and Spark.
We use three simple synthetic applications to have a deep understanding of the 
underlying behavior of the applications with simple I/O patterns, namely
\textit{Increment}, \textit{Multi-Increment}, and \textit{Histogram}.
Also, two neuroimaging applications are used to study more realistic 
applications, namely \textit{Kmeans} and \textit{BIDS App Example}\HL{{\cite{gorgolewski2017bids}}}.

@@

\HL{The scripts used for our experiements and the results generated are available on
	GitHub at: \mbox{\href{https://github.com/mathdugre/paper-big-data-engines}{https://github.com/mathdugre/paper-big-data-engines/}}.}

\subsubsection{Increment}
We adapted the increment application used in~\cite{8752675}.
This synthetic application reads blocks of the BigBrain from Lustre and
simulates computation by sleeping for a specified period (Fig.~\ref{fig:graph-increment}). To simulate
intermediate results, we repeat \HL{a {\SI{1}{\second}} sleep delay during each incrementation}.
We prevent data caching of the blocks by incrementing their voxels
value by one after each sleep operation. Finally, we write the resulting
NIfTI image back to Lustre. This application allows us to study the engines
when their inputs are processed independently. The map-only pattern
mimics the processing of multiple independent subjects in parallel.
% \begin{figure}[!hb]
% 	\centering
% 	\includegraphics[height=0.5\columnwidth,
% 	angle=0]{figures/increment.png}
% 	\caption{Task graph for Incrementation with 3 iterations and 3 BigBrain blocks.}
% 	\label{fig:graph-increment}
% \end{figure}

@@

\subsubsection{Kmeans}
For our fourth application, we apply Kmeans clustering to the voxel
intensities of the BigBrain image. We set the number of clusters to 3. The
application starts by reading the image blocks, combining all voxels in a
1-D array, and choosing initial centroids using the min, max, and
intermediate values. It assigns each voxel to its closest centroid and
updates each centroid with the average of its assigned voxels. It repeats
the assignment and update steps for a configurable number of iterations.
Finally, the voxels of the image blocks are classified and written back to
the file system. Updating the centroids involves substantial data
communication between the workers.
The computation to find the closest centroids for each voxels scales with the number of centroids.
It can be computed in parallel at the cost of higher memory or sequentially to reduce memory usage.
We decided to compute the nearest centroids in parallel since we use only three centroids.

\HL{Due to resources limitation on our infrastructure, the Kmeans experiments
	are performed with only a sample of half the BigBrain dataset ({\SI{303}{\gibi\byte}}).}

\subsubsection{BIDS App Example}
Our fifth application is the BIDS App Example, a neuroimaging pipeline to
measure brain volume from MRIs. For this application, we use the CoRR
dataset. The application extracts the brain volume of each participant,
then computes the average for each group of participants. Unlike the other
applications, the BIDS App Example is a command-line executed in a Docker image
(bids/example on DockerHub). We converted the Docker image to a Singularity
image for use in HPC environments, using
\href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}

\subsection{Experiments}
Table~\ref{table:parameters} shows the parameters that we varied
throughout the experiments. We varied (1) the number of nodes to assess
the scalability of the engines' scheduler, (2) the number of files for the BigBrain
in Increment, and Histogram to measure the
effect of different I/O patterns and parallelization degrees, and (3) the
number of iterations to evaluate the effect of the number of tasks.
It should be noted that increasing the number of iterations for a given sleep 
delay also increases the total compute time of an application.

@@

\section{Results}
\subsection{Increment}
Figure \ref{fig:increment_worker} depicts the total execution time spent in each code segment of the \textit{Increment} application.
The bars are broken down into Overhead, Read, Write, and other application-specific functions.
The mean of each bar is calculated from ten repetitions, with the error bar representing the standard deviation.
As expected, the compute time, from \textit{Increment}, remains the same when we vary the number of nodes.
However, the \textit{Read} and \textit{Write} time increase with the number of nodes for both engines.
This is explained by an I/O bottleneck on the Lustre file system.

From Figure \ref{fig:increment_worker}, we also observe that Dask's overhead is higher than Spark's.
While the overhead of both engines increases with the number of nodes, Dask's overhead is significantly worse when scaling.
\HL{Although Dask has a higher overhead, the total time for both engines is overall significantly similar.}
This is due to a tradeoff between disk bandwidth and compute/overhead time.
On the one hand, multiple threads access Lustre concurrently, thus sharing the disk bandwidth resulting in longer I/O time.
On the other hand, when most threads are either idle or performing computations, it reduces the contention on the disk bandwidth resulting in shorter I/O time.
\HL{The only significant difference in total time is when two nodes are used.
	This results in an approximately 3\% total runtime difference.}
\begin{figure}[!h]
	\centering
	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_increment_worker.jpg}
	\caption{Increment: varying nodes -- 5000 files, 5 iterations}
	\label{fig:increment_worker}
\end{figure}

@@

Figure \ref{fig:increment_itr} also shows that the the total I/O time decreases as total computation time increases.
This is explained by an I/O bottleneck on the Lustre file system.
When threads are busy with computations, it reduces the contention on Lustre, thus improving I/O speed for individual accesses.
\begin{figure}[!h]
	\centering
	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_increment_itr.jpg}
	\caption{Increment: varying iterations -- 4 nodes, 5000 files}
	\label{fig:increment_itr}
\end{figure}

@@

From Figure \ref{fig:increment_block} we also observe that Dask has significantly higher overhead than Spark.
However, the total execution time of the application is \HL{significantly equivalent} due to the tradeoff between I/O and compute/overhead time.
\begin{figure}[!h]
	\centering
	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_increment_block.jpg}
	\caption{Increment: varying \# of files -- 4 nodes, 5 iterations}
	\label{fig:increment_block}
\end{figure}

\subsection{Histogram}
Figure \ref{fig:histogram_worker} depicts the execution time for the \textit{Histogram} application when varying the number of nodes.
Overall, \HL{the execution time for Spark is significantly lower than for Dask by up to 5\%}, with the difference coming from a shorter execution time for \textit{Calculate\_histogram}.
This is unexpected as the code base for both engines is the same, with only the calling APIs differing.
It is most likely due to the use of Dask Bags, which are slower than the other Dask APIs.
As with previous experiments, the overhead for Spark is less and scales better, although compensated by a longer total I/O time.
\begin{figure}[!h]
	\centering
	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_histogram_worker.jpg}
	\caption{Histogram: varying nodes -- 5000 files}
	\label{fig:histogram_worker}
\end{figure}

>>>>

\abstract[Summary]{
	The general increase in data size and data sharing motivates the adoption of Big Data strategies in several scientific disciplines.
	However, while several options are available, no particular guidelines exist for selecting a Big Data engine.
	In this paper, we compare the runtime performance of two popular Big Data engines with
	Python APIs, Apache Spark, and Dask, in processing neuroimaging pipelines.
	Our experiments use three synthetic data-intensive applications to process
	the \SI{606}{\gibi\byte} BigBrain image and an actual pipeline to process data
	from thousands of anatomical images.
	We benchmark these applications on a dedicated HPC cluster running the Lustre file system while using varying
	combinations of the number of nodes, file size, and task duration.
	Our results show that although there are slight differences between Dask and Spark,
	the performance of the engines is comparable for data-intensive applications.
	However, Spark requires more memory than Dask, which can lead to 
	slower runtime depending on configuration and infrastructure.
	In general, the limiting factor was the data transfer time.
	While both engines are suitable for neuroimaging, more efforts need to be put
	to reduce the data transfer time and the memory footprint of applications.
}

@@

\section{Introduction}
The rise in data sharing coupled with improved data collection technologies leads neuroimaging into the Big Data era\cite{ALFAROALMAGRO2018400, van2014human, ConpPortal}.
While standard neuroimaging workflow engines, such as Nipype\cite{Nipype:11}, are well rounded to process standard compute-intensive pipelines,
they lack Big Data strategies (i.e., in-memory computing, data locality, and lazy-evaluation) to improve the performance of increasingly prevalent data-intensive pipelines.
A previous study\cite{8752675} noted the importance of in-memory computing and data locality to improve the performance of data-intensive pipelines.
This work studies how performance benchmarks generalize across Big Data engines in an HPC environment.
In particular, we study in-depth the performance difference between Dask\cite{Dask:15} and Apache Spark\cite{Spark:16} for their suitability to process neuroimaging pipelines.
Our objective is to assess whether Spark or Dask has substantial performance benefits in processing data-intensive HPC pipelines.

Spark and Dask both provide in-memory computing, data locality, and lazy-evaluation, typical for Big Data engines.
The scheduler for both engines operates dynamically, which benefits applications with runtime unknown ahead of time\cite{Dask:15}.
They also provide rich high-level APIs and support a variety of schedulers and deployment environments, such as Mesos\cite{hindman2011mesos}, YARN\cite{vavilapalli2013apache}, Kubernetes, and HPC clusters.
Although sharing similarities, these engines have differences.

First and foremost, Spark is written in Scala while Dask is in Python.
The popularity of Python in the scientific community arguably gives an edge to Dask due to the serialization cost between Python and Scala.
However, if not careful, the Python Global Interpreter Lock (GIL) can significantly reduce the parallelism of an application in some cases.
The difference in programming languages also provides different benefits.
On the one hand, as part of the Scipy ecosystem, Dask provides almost transparent integration with APIs such as Numpy arrays, Pandas dataframe, or RAPIDS GPU accelerated code framework.
On the other hand, Spark's Java, Scala, R, and Python APIs efficiently parallelize pipelines with minimal performance loss due to the Java Virtual Machine.
Our study is restrained to performance, although we recognize that other factors affect the choice of an engine in practice.

While laptops or workstations can be sufficient for some applications, data-intensive neuroimaging pipelines often require large infrastructures.
This paper focuses on the performance of Big Data engines in HPCs environments.
We installed the Slurm scheduler and Lustre file system~\cite{Braam2019TheLS} on a dedicated cluster to mimic a typical HPC system.
Lustre is a reference file system for HPC environments, used in over 60 TOP100 supercomputers~{\cite{OpenSFS-lustre}}, offering parallel file accesses.

In a previous study\cite{8943502} we found no significant performance difference between Apache Spark and Dask for data-intensive applications.
However, using the network file system (NFS) led to an I/O bottleneck limiting application performance.
Another study~\cite{8588652} also reported network and I/O bottlenecks when concurrency increased.
This same study found that the startup overhead time for Spark was more prominent than for Dask.
On the contrary, the work in~\cite{Mehta:17} claims that the startup overhead is more significant for Dask than for Spark.
This contradiction suggests that the startup overhead might vary depending on cluster configuration.
The latter study processed a \SI{100}{\giga\byte} dataset using a neuroimaging pipeline and found Dask to be up to 14\% faster than Spark due to ``more efficient pipelining" and serialization time to Python.
On the contrary, the work in~\cite{10.1145/3225058.3225128} shows that with large datasets, Spark provides better speedup than Dask.
However, that study used the Dask Bag API, which has performance limitations compared to other Dask APIs.

The above studies compared the performance of the engine at a high level.
In comparison, our study provides a detailed analysis of the performance differences
and similarities between Dask and Spark for data-intensive applications executed on HPC systems.

@@

\section{Background}
\subsection{Lustre}
The Lustre file system is a parallel file system of reference in HPC environments.
Lustre offers a POSIX-compliant interface with the file system.
Installed in some of the most powerful supercomputers, the design of Lustre allows
scaling to thousands of nodes, petabytes of storage, and hundreds of gigabits per second of I/O throughput.

At the core of Lustre are the \textit{management servers} (MGSs) and \textit{metadata server} (MDSs) in charge of the file operations and metadata,
the \textit{object storage targets} (OSTs) store file data, and the \textit{object storage servers} (OSSs) handle the OSTs.
Lustre clients make requests to the MGSs to access files on the OSTs.
To handle failures in the system, the MGSs, MDSs, and OSTs have failover capabilities.
InfiniBand, Ethernet, or a combination of both can be used to interconnect these different components.

While other filesystems such as HDFS or cloud storage could be considered, we chose Lustre to represent an HPC environment best.
HDFS uses permanent local storage on compute node, which is usually not feasible in HPC.
Moreover, HPC nodes rarely have access to the internet, limiting cloud storage use.

\subsection{Dask}
Dask is a Python-based Big Data engine with growing popularity in the scientific Python ecosystem.
Dask was designed with data locality and in-memory computing in mind to mitigate the data transfer bottleneck in Big Data workflows.
Data locality, popularized by Map-Reduce~\cite{dean2008mapreduce}, schedules tasks where the data reside.
In-memory computing minimizes the overhead of transferring data to disks by keeping data in memory.
Dask uses lazy evaluation to reduce unnecessary communication and computation.
The engine builds a dynamic graph before execution, allowing it to determine which task to compute.
Dask workflows can further reduce data transfers by leveraging multithreading whenever the Python GIL does not restrict it.
It achieves fault tolerance by recording data lineage, the sequence of operations used to modify the initial data.

Dask offers five data structures:
\href{https://docs.dask.org/en/latest/array.html}{Array},
\href{https://docs.dask.org/en/latest/bag.html}{Bag},
\href{https://docs.dask.org/en/latest/dataframe.html}{DataFrame},
\href{https://docs.dask.org/en/latest/delayed.html}{Delayed},
and \href{https://docs.dask.org/en/latest/futures.html}{Futures}.
Arrays offer a clone of NumPy API for distributed processing of large arrays.
Bags are a distributed collection of Python objects that offers a programming abstraction similar to \href{https://toolz.readthedocs.io/en/latest/}{PyToolz}.
Dataframes are a parallel composition of \href{https://pandas.pydata.org/}{Pandas} Dataframes used to process a large amount of tabular data.
Dask Delayed offers an API for distributing arbitrary functions that do not fit the above frameworks.
Lastly, Dask Futures can also execute arbitrary functions; however, it launches computation immediately rather than lazily.
Dask allows users to install only required components making it lightweight.

@@

The Dask engine is compatible with multiple distributed schedulers, including YARN and Mesos.
It also provides its scheduler, \textit{Dask Distributed scheduler}.
We chose to use Dask Distributed scheduler to keep the environments similar between Dask and Spark.

@@

In the Spark Standalone scheduler, a \textit{leader} (a.k.a master) coordinates the resource provisioned by \textit{workers} in the cluster.
A \textit{driver}  process receives jobs from clients and requests workers from the leader.
Jobs are divided into stages to execute onto workers.
Each operation in a stage is represented by a high-level task in the computation graph.
Like Dask, Spark Standalone scheduler uses a LIFO policy to schedule tasks.
Spark Standalone has two execution modes: (1) the client mode, where the driver process runs in a dedicated process,
and (2) the cluster mode, where the driver runs within a worker process.
Our experiments use the client mode since cluster mode is not available in PySpark.

@@

Python is a standard programming language in the scientific community, offering numerous data processing libraries.
While serialization from Python to Java, an operation required when using Spark's Python API, creates overhead, we found it minimal in a previous study~\cite{8943502}.
We focus on PySpark API to have a similar environment between the different engines and  its suitability for neuroimaging.

\section{Methods}
\subsection{Infrastructure}
We used the ``slashbin" cluster at Concordia University.
The cluster has eight compute nodes, four storage nodes, one login node, and one Lustre metadata node.
Each compute node has two 16-core Intel(R) Xeon(R) Gold 6130 CPU @ 2.10GHz,
\SI{250}{\gibi\byte} of RAM, \SI{126}{\gibi\byte} of tmpfs,
% 6 $\times$ SSDs of \SI{447}{\gibi\byte} each with the XFS file system (no RAID enabled),
CentOS~8.1 and Linux kernel \textit{4.18.0-240.1.1.el8\_lustre.x86\_64}.
The compute and storage nodes are inter-connected by {\SI{10}{\giga\bit\per\second}} Ethernet link through a Mellanox MSN2410 switch.
The login node, hosting an NFS server, is connected to the compute nodes through the same switch, with {\SI{10}{\giga\bit\per\second}} Ethernet.

We configured Lustre to have 1 MGS and 4 OSS.
Each OSS has 11 OSTs (disks) of \SI{8.8}{\tebi\byte}.
The OSTs are configured to have a maximum of \SI{1000}{\mebi\byte} of dirty data written into the client page cache.
Both the MDT and OSTs are configured to a maximum of 64 concurrent RPCs in flight.
A timeout of \SI{100}{\second} is set for Lustre, and OSTs checksum are disabled.
Lustre \textit{2.14.0-1.el8} is installed on both the Lustre clients and servers.
We used the default inode configuration for Lustre.

Our system uses similar hardware as Compute Canada's B\'eluga cluster --- a representative cluster available to researchers in Canada --- and replicates as
close as possible its configuration.
This lets us compare Spark and Dask in the condition of a realistic
HPC system.

Both Spark and Dask are configured to have eight worker processes, each with eight threads.
Each worker was allocated \SI{31.5}{\gibi\byte} of memory.
For Spark, the JVM heap space uses 10\% of workers' memory, and the location of the log is set on an NFS-mounted directory.
The worker workspace for both engines is located on Lustre due to the limited fast storage space on the worker nodes.
Other Dask configurations are left to default.
The Spark network timeout is set to \SI{600}{\second} and the executor heartbeat to \SI{120}{\second}.
A new Dask or Spark cluster was started and torn down for each experiment.
Dask \textit{2021.7.0} and Spark \textit{3.1.2} were used.

\subsection{Dataset}
We used BigBrain\cite{Amunts:13}, a 3-D image of the human brain with voxel
intensities ranging from 0 to 65,535. We converted the blocks into the
NIfTI format, a popular format in neuroimaging. We left the NIfTI blocks
uncompressed, resulting in a total data size of \SI{606}{\gibi\byte}. To
evaluate the effect of block size, we split the image into 1000, 2500,
and 5000 files of \SI{606}{\mebi\byte}, \SI{242.4}{\mebi\byte}, and
\SI{121.2}{\mebi\byte}, respectively.
We used the \href{https://github.com/big-data-lab-team/sam}{sam} library to split the image.

@@

\subsection{Applications}
This section describes the applications used to benchmark Dask and Spark.
We use three simple synthetic applications to have a deep understanding of the 
underlying behavior of the applications with simple I/O patterns, namely
\textit{Increment}, \textit{Multi-Increment}, and \textit{Histogram}.
Also, two neuroimaging applications are used to study more realistic 
applications, namely \textit{Kmeans} and \textit{BIDS App Example}{\cite{gorgolewski2017bids}}.

@@

The scripts used for our experiments and the results generated are available on
GitHub at: \mbox{\href{https://github.com/mathdugre/paper-big-data-engines}{https://github.com/mathdugre/paper-big-data-engines/}}.

\subsubsection{Increment}
We adapted the increment application used in~\cite{8752675}.
This synthetic application reads blocks of the BigBrain from Lustre and
simulates computation by sleeping for a specified period (Fig.~\ref{fig:graph-increment}). To simulate
intermediate results, we repeat a {\SI{1}{\second} sleep delay during each incrementation}.
We prevent data caching of the blocks by incrementing their voxels
value by one after each sleep operation. Finally, we write the resulting
NIfTI image back to Lustre. This application allows us to study the engines
when their inputs are processed independently. The map-only pattern
mimics the processing of multiple independent subjects in parallel.

@@

\subsubsection{Kmeans}
For our fourth application, we apply Kmeans clustering to the voxel
intensities of the BigBrain image. We set the number of clusters to 3. The
application starts by reading the image blocks, combining all voxels in a
1-D array, and choosing initial centroids using the min, max, and
intermediate values. It assigns each voxel to its closest centroid and
updates each centroid with the average of its assigned voxels. It repeats
the assignment and updates steps for a configurable number of iterations.
Finally, the voxels of the image blocks are classified and written back to
the file system. Updating the centroids involves substantial data
communication between the workers.
The computation to find the closest centroids for each voxels scales with the number of centroids.
It can be computed in parallel at the cost of higher memory or sequentially to reduce memory usage.
We decided to compute the nearest centroids in parallel since we use only three centroids.
Due to resources limitation on our infrastructure, the Kmeans experiments
are performed with only half of the BigBrain dataset ({\SI{303}{\gibi\byte}}).

\subsubsection{BIDS App Example}
Our fifth application is the BIDS App Example, a neuroimaging pipeline to
measure brain volume from MRIs. For this application, we use the CoRR
dataset. The application extracts the brain volume of each participant
then computes the average for each group of participants. Unlike the other
applications, the BIDS App Example is a command-line executed in a Docker image
(bids/example on DockerHub). We converted the Docker image to a Singularity
image for use in HPC environments, using
\href{https://hub.docker.com/r/singularityware/docker2singularity/tags/}{docker2singularity}

\subsection{Experiments}
Table~\ref{table:parameters} shows the parameters that we varied
throughout the experiments. We varied (1) the number of nodes to assess
the scalability of the engines' scheduler, (2) the number of files for the BigBrain
in Increment and Histogram to measure the
effect of different I/O patterns and parallelization degrees, and (3) the
number of iterations to evaluate the effect of the number of tasks.
It should be noted that increasing the number of iterations for a given sleep 
delay also increases the total compute time.

@@

\section{Results}
\subsection{Increment}
Figure \ref{fig:increment_worker} depicts the total execution time spent in each code segment of the \textit{Increment} application.
The bars are broken down into Overhead, Read, Write, and other application-specific functions.
The mean of each bar is calculated from ten repetitions, with the error bar representing the standard deviation.
As expected, the compute time, from \textit{Increment}, remains the same when we vary the number of nodes.
However, the \textit{Read} and \textit{Write} time increases with the number of nodes for both engines.
This is explained by an I/O bottleneck on the Lustre file system.

From Figure \ref{fig:increment_worker}, we also observe that Dask's overhead is higher than Spark's.
While the overhead of both engines increases with the number of nodes, Dask's overhead is significantly worse when scaling.
Although Dask has a higher overhead, the total time for both engines is overall similar.
This is due to a tradeoff between disk bandwidth and compute/overhead time.
On the one hand, multiple threads access Lustre concurrently, thus sharing the disk bandwidth resulting in longer I/O time.
On the other hand, when most threads are either idle or performing computations, it reduces the contention on the disk bandwidth resulting in shorter I/O time.
The only significant difference in total time is when two nodes are used.
This results in an approximately 3\% total runtime difference.
\begin{figure}[!h]
	\centering
	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_increment_worker.jpg}
	\caption{Increment: varying nodes -- 5000 files, 5 iterations}
	\label{fig:increment_worker}
\end{figure}

@@

Figure \ref{fig:increment_itr} also shows that the total I/O time decreases as total computation time increases.
This is explained by an I/O bottleneck on the Lustre file system.
When threads are busy with computations, it reduces the contention on Lustre, thus improving I/O speed for individual accesses.
\begin{figure}[!h]
	\centering
	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_increment_itr.jpg}
	\caption{Increment: varying iterations -- 4 nodes, 5000 files}
	\label{fig:increment_itr}
\end{figure}

@@

From Figure \ref{fig:increment_block} we also observe that Dask has significantly higher overhead than Spark.
However, the total execution time of the application is equivalent due to the tradeoff between I/O and compute/overhead time.
\begin{figure}[!h]
	\centering
	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_increment_block.jpg}
	\caption{Increment: varying \# of files -- 4 nodes, 5 iterations}
	\label{fig:increment_block}
\end{figure}

\subsection{Histogram}
Figure \ref{fig:histogram_worker} depicts the execution time for the \textit{Histogram} application when varying the number of nodes.
Overall, the execution time for Spark is significantly lower than for Dask by up to 5\%, with the difference coming from a shorter execution time for \textit{Calculate\_histogram}.
This is unexpected as the code base for both engines is the same, with only the calling APIs differing.
It is most likely due to the use of Dask Bags, which are slower than the other Dask APIs.
As with previous experiments, the overhead for Spark is less and scales better, although compensated by a longer total I/O time.
\begin{figure}[!h]
	\centering
	\includegraphics[clip,width=0.75\columnwidth]{figures/stacked_histogram_worker.jpg}
	\caption{Histogram: varying nodes -- 5000 files}
	\label{fig:histogram_worker}
\end{figure}
